Renal IV

### 2017-06-08 renal4.rmd
fixed some typos

-  Learn bladder Ca
    -  Smoking is major risk factor. Can present as painless haematuria
-  Major mech of stress incontinence: *urethral hypermobility*
-  *Proteus mirabilis* is a gram(-) rod that leads to highly alkaline urine
-  Medication associated with acute urinary retention: decongenstant
-  Enuresis: bedwetting during sleep in child >6
-  Pharm Mx of BPH: $\alpha$-adrenergic receptor antagonists
-  Clavulanic acid contains beta-lactam, gets preferentially broken down by resistant bacteria. Therefore penicillin and cephalosporins can have an effect.
-  Receptors
    -  M~3~: ACh from parasympathetic pelvic nerve to contract detrusor
    -  B~3~: hypogastric N (sym): noradrenaline relaxation of detrusor
    -  $\alpha_{1}$ in the int sphincter: hypogastric N: contraction of ext sphincter
    -  Nicotinic in the ext sphincter: somatic from pudendal nerve to contract it
    -  Also sensory N AFFERENT in detrusor that is stimulated when bladder is stretched
-  Calculi:
    -  Struvite likely to be from infn

@@

-  Somatic vs visceral: things affecting visceral peritoneum would be very sharp due to lots of nerves
-  Nociceptive (pain due to nociceptors detecting tissue damage. Cna be blocked with NSAIDS, opiods; sharp, burning) vs neuropathic (irritated/diseased nerves e.g. peripheral/autoimmune/diabetic neuropathy or obstructions; described as tingling, burning, pins and needles)
-  Types of abdo pain: epigastric (foregut), umbilical (midgut), lower abdo (hindgut)?!?!: most likely to be hindgut pain (2/3 of transverse colon and down)

@@

## Trogger 2

@@

Risks of cathetirsation: urosepsis (septicaemia due to infection in the genitourinary system)

@@

Medications interfering with micturition such as NSAIDS, decongestants for URTI

>>>>

-  Learn bladder Ca
    -  Smoking is major risk factor. Can present as painless haematuria
-  Major mech of stress incontinence: *urethral hypermobility*
-  *Proteus mirabilis* is a gram(-) rod that leads to highly alkaline urine
-  Medication associated with acute urinary retention: decongenstant
-  Enuresis: bedwetting during sleep in child >6
-  Pharm Mx of BPH: $\alpha$-adrenergic receptor antagonists
-  Clavulanic acid contains beta-lactam, gets preferentially broken down by resistant bacteria. Therefore penicillin and cephalosporins can have an effect.
-  Receptors
    -  M~3~: ACh from parasympathetic pelvic nerve to contract detrusor
    -  B~3~: hypogastric N (sym): noradrenaline relaxation of detrusor
    -  $\alpha_{1}$ in the int sphincter: hypogastric N: contraction of ext sphincter
    -  Nicotinic in the ext sphincter: somatic from pudendal nerve to contract it
    -  Also sensory N AFFERENT in detrusor that is stimulated when bladder is stretched
-  Calculi:
    -  Struvite likely to be from infn (such as *Proteus*)

@@

-  Somatic vs visceral: things affecting visceral peritoneum would be very sharp due to lots of nerves
-  Nociceptive (pain due to nociceptors detecting tissue damage. Can be blocked with NSAIDS, opiods; sharp, burning) vs neuropathic (irritated/diseased nerves e.g. peripheral/autoimmune/diabetic neuropathy or obstructions; described as tingling, burning, pins and needles)
-  Types of abdo pain: epigastric (foregut), umbilical (midgut), lower abdo (hindgut)?!?!: most likely to be hindgut pain (2/3 of transverse colon and down)

@@

@@

Risks of catheterisation: urosepsis (septicaemia due to infection in the genitourinary system)

@@

Medications interfering with micturition such as NSAIDS, decongestants for URTI.
